---
title: "Introduction to r 1"
author: "emma"
date: "6/13/2020"
output: pdf_document
---
#options(repos='http://cran.rstudio.com/')

```{r}
# set directory
setwd("C:/Users/USER/Desktop/JUPYTER_NOTEBOOK/A_MYTUTORIALS/MYR")
#Lists the current working directory
getwd()
# objects in the current workspace
ls()
options(digits=3)
a<-2.33333444
a
```

# Vectors

creating an empty vector and populate it with 8 and 9
```{r}
x <- vector()
x[1]<-8
x[2]<-9
x
```



```{r}
numeric <- c(1, 2.5, 4.5)
numeric
# L distinguishes doubles from integers
integer <- c(1L, 6L)
typeof(integer)
```

```{r}
v <- c(34, 18, 23)
length(v)
mode(v)
```

All elements of a vector must belong to the same mode. If that is not true,
R will force it by type coercion
```{r}
v <- c(34, 18, 23,'a')
length(v)
mode(v)
v
```
modulo 
```{r}
9%%2
```


```{r}
# Use T and F or TRUE and FALSE to create logical vectors
logical <- c(T, FALSE, TRUE, F)
typeof(logical)
character <- c("these are called characters","or", "some strings")
character
a<-1:5 #sequence from 1 to 5
a
```

## factors
 Factors are used to represent categorical data. Factors can be ordered or unordered and are an important class for statistical analysis and for plotting.Factors are stored as integers,and have labels associated with these unique integers. While factors look (and often behave) 
```{r}
ap<-c('red','gren','black')
ap
fact.ap<-factor(ap)
fact.ap
nlevels(fact.ap)
```

```{r}

rank<-c('low','medium','high')
fack<-factor(rank,levels=c('medium','low','high'),order=T)
fack
min(fack)
max(fack)

rank<-c('low','medium','high')
fac<-factor(rank)
fac
rank<-factor(rank,levels=c('low','medium','high'),ordered=TRUE)

as.numeric(rank)
min(rank)

```


```{r}


data = c(1,2,2,3,1,2,3,3,1,2,3,3,1)
fdata = factor(data)
fdata

rdata = factor(data,labels=c("I","II","III"))
rdata
#To convert the default factor fdata to roman numerals, we use the assignment form of the levels function:
levels(fdata) = c('I','II','III')
table(fdata)
```



```{r}
## CREATING ARRAYS
print('arrays')
array(c(1,2,3,4),dim = c(2,2))
A<-array(c(1,'b','c'),dim = c(2,2,3))
print(A)
is.double(A)
is.recursive(A)
```


##  CREATING A MATRIX
```{r}
print("go")
a <-matrix(c(1,2,3,4),nrow = 2,byrow = T)
a
# matrix is by default column
matrix(c(1,2,3,4),nrow = 2)
# testing to see if a is a matrix
is.matrix(a)
t(a)#finding the transpose of a matrix
matrix(c(1,2,3,4),ncol = 2)

```



```{r}
f<-matrix(c(3,5,6,8),nrow=2,byrow = T)
f
print('multiplication of f by f')
f%*%f # matrix multiplication
print('inverse of f')
solve(f)# inverse of a matrix
t(f)# transpose of a matrix
```


```{r}
x<-c(1:12)
x
#reshaping x into a 3 by 4 matrix
dim(x)<-c(3,4)
x
# number of rows
nrow(x)
#number of columns
ncol(x)
rownames(x)<-c('row1','row2','row3')
colnames(x)<-c('col1','col2','col3','col4')
x
```




# LSIT
```{r}
x <- list(1:3, c('a','b'), c(T, F, T), c(2.3, 5.9,3))
#Attributes can be accessed individually with attr() or all at once (as a list) with attributes().
attr(x,'my_attributes')<-'this is a list'

x
is.recursive(x)
str(x)

```
```{r}
x<-c(1:4)
names(x)<-c('a','b','c','d')
x

xx<-setNames(c(1:3),c('col1','col2','col3'))
xx
```
## CREATING DATA FRMAES

```{r}
a<-data.frame(name=c('ama','kofi','yaw'),age=c(24,9,10),sex=c('female','male','male'))
a
rownames(a)<-c(1,2,3)
a
```

```{r}
d<-c(1,2,3,4,4,5)
f<-c(3,4,5)
cat('d values are',d ,'\n','f values are',f)

dd<-c(1,2,3,4,4,5)
ff<-c('a','b')
cat('dd values are',class(dd) ,'\n','ff values are',class(ff))


```



```{r}
print('object in the workspace containing fac')
ls(pattern='fa')

print("ok")
ls(all.names = T)

```

# modulus
# to print the remainder
```{r}
8%%2
5%%2
```



```{r}
# relational operators
d<-c(5,2,3,5)
f<-c(3,5,67,8)
d>f
d<=f
d==f

```

```{r}
# LOGICAL OPERATORS
d<-c(1,FALSE,3,5)
f<-c(3,5,67,8)
d&f # & AND operator
print('not d')
!d # ! not operator
d|f # | OR operator
d<-c(0,FALSE,3,5)# ZERO IS CONSIDER AS FALSE
f<-c(3,5,67,8)
d&f
k<-c(T,F)
k
!k
```


```{r}
# SEQUNCE
seq(1:4)
seq(1,4,by=0.5)
seq(1,5,length=9)
```

```{r}
#repeat function
rep(c(1,2),c(4,2))#repeat 1 4times and 2 2times
rep(1,2)#repeat 1 2times
rep(c(2,3),4)#repeat 2,3 4times

```
```{r}
x<-c(1,2,3,4,NA)
x

mean(x)
mean(x,na.rm =T)
na.omit(x)
mean(na.omit(x))
```
```{r}
library(RColorBrewer)
n <- 9
mycolors <- brewer.pal(n, "Set3")
barplot(rep(1,n), col=mycolors)
```


#Combining data frames
You can combine data frames using cbind() and rbind()
```{r}
dosage<-c(20,30,40,45,60)
Res_Drug_A<-c(20,30,40,45,60)

Res_Drug_B<-c(15,18,25,31,40)
rowsie<-c(28,90,40)
# conbiming columnwise
data<-cbind(data.frame(dosage),data.frame(Res_Drug_A),data.frame(Res_Drug_B))
# rowwise concatenation
data<-rbind(data,rowsie)
data
```


```{r}
plot(Res_Drug_A,Res_Drug_B,type = "b",pch=20,lty=17,cex=2.5,lwd=3,col=c("blue","green","red","yellow","orange"),col.lab=(2),bg=col)
```



```{r}
runif(1)

# Get a vector of 4 numbers
runif(4)

# Get a vector of 3 numbers from 0 to 100
runif(3, min=0, max=100)


# Get 3 integers from 0 to 100
# Use max=101 because it will never actually equal 101
floor(runif(3, min=0, max=101))
```



```{r}
# This will do the same thing
sample(1:100, 3, replace=TRUE)


# To generate integers WITHOUT replacement:
sample(1:100, 3, replace=FALSE)

#To generate numbers from a normal distribution, use rnorm(). By default the mean is 0 and the standard deviation is 1.
rnorm(4)


# Use a different mean and standard deviation
rnorm(4, mean=50, sd=10)

# To check that the distribution looks right, make a histogram of the numbers
x <- rnorm(400, mean=50, sd=10)
hist(x)
```




```{r}
# generating uniform random numbers
runif(1)

# Get a vector of 4 numbers
runif(4)

# Get a vector of 3 numbers from 0 to 100
runif(3, min=0, max=100)

# Get 3 integers from 0 to 100
# Use max=101 because it will never actually equal 101
floor(runif(3, min=0, max=101))

# This will do the same thing
sample(1:100, 3, replace=TRUE)
sample(1:13,3 ,replace = T)


# To generate integers WITHOUT replacement:
sample(1:100, 3, replace=FALSE)
sample(1:3,3)# by default is sample without replacement

# RANDOM NUMBERS FROM NORMAL DISTRIBUTION

#To generate numbers from a normal distribution, use rnorm().
#By default the mean is 0 and the standard deviation is 1.

rnorm(4)

# Use a different mean and standard deviation
rnorm(4, mean=50, sd=10)

# To check that the distribution looks right, make a histogram of the numbers
x <- rnorm(400, mean=50, sd=10)
hist(x)
#par(mfrow=c(1,3))
```











