---
title: "Untitled"
author: "emma"
date: "6/25/2020"
output: pdf_document
---


```{r}
x <- c(2,9, 4.2,5,6,8,10.2)
x[1]
x[c(3,2)]
x[c(1:4)]
x[order(x)]
order(x)# print the position of the values in x in ascending order
#omit the first and second value
x[-c(1, 2)]
x[x>4]
x[]#Nothing returns the original vector
```

```{r}
a <- matrix(1:9, nrow = 3)
colnames(a)<-c('col1','col2','col3')
a
a[1:3,c('col1')]

a[,c('col2')]
```
```{r}
a<-outer(1:3,1:3,FUN = "paste",sep=",")
a
```

```{r}
a[1:2,1:2]
```
```{r}
x <- outer(1:3, 1:3, FUN = "*")
x
```

```{r}
x[c(1:2),c(2:3)]# row 1,2 and column 2,3
```

assigning 21 to row 1 col 2, 22 to row 1 col 3 , 30 to row 2 col 2, 40 to row 2 col 

```{r}
x[c(1:2),c(2:3)]<-c(21,22,30,40)
cat(" print all values of x without column 2 \n")
x[,-2]#
cat(" print all values of x without column 1 and 2 \n")
x[,c(-1,-2)]
```
subsetting a dataframe
```{r}
df<-data.frame(name=c('ama','kofi','yaw','emma'),age=c(24,20,9,10),
              sex=c('female','male','male','female'))
df$name
```

```{r}
df[c('age','sex')]
```
Select 2 random rows without replacement
```{r}
df[sample(nrow(df),2),]
```
Select 4 random rows with replacement
```{r}
df[sample(nrow(df),5,replace = T),]
```
```{r}
reorder_col<-df[sample(nrow(df),5,replace = T),3:1]
reorder_col
```
```{r}

mtcars<-mtcars
# structure of the Motor Trend Car Road Tests
str(mtcars)
# first few elements of the data
head(mtcars)
# shape of the data
dim(mtcars)

```
```{r}
summary(mtcars$mpg)
plot(mtcars$mpg, mtcars$disp,main = 'mpg vrs disp')
plot(mtcars$mpg, mtcars$wt,main = 'mpg vrs wt')
```
The attach() function adds the data frame to the R search path
Detach Objects from the Search Path
```{r}
attach(mtcars)
summary(mpg)
plot(mpg, disp,main = 'mpg vrs disp')
plot(mpg,wt,main = 'mpg vrs wt')
detach(mtcars)
```
subsetting instances where mpg is less than 15
```{r}
mtcars[mtcars$mpg <15,]
```
subsetting instances where mpg is less than 15 and ls disp<400
```{r}
mtcars[mtcars$mpg <15 & mtcars$disp <400,]
```

subsetting instances where mpg is less than 15 OR disp<400
```{r}
mtcars[mtcars$mpg <15 || mtcars$disp <400,]

```

The limitation of the with() function is that assignments exist only within the
function brackets
```{r}
with(mtcars,{
  s<-summary(mpg)
  s
})
#su # does not exist outside the brackets
```
for s to exist outside the brackets use <<
```{r}
with(mtcars,{
  s<<-summary(mpg)
})
s
```

