---
title: "INTRODUCTION TO R"
author: "ARIZE"
date: ""
output: pdf_document
---


```{r}

```


```{r}
# switch
x<-switch(1,"0ne","two","three")
x

y<-switch(3,"0ne","two","three")
y
```

```{r}
number1 <- 30
number2 <- 20
operator <- readline(prompt="Please enter any ARITHMETIC OPERATOR You wish!: ")
 
switch(operator,
       "+" = print(paste("Addition of two numbers is: ", number1 + number2)),
       "-" = print(paste("Subtraction of two numbers is: ", number1 - number2)),
       "*" = print(paste("Multiplication of two numbers is: ", number1 * number2)),
       "^" = print(paste("Exponent of two numbers is: ", number1 ^ number2)),
       "/" = print(paste("Division of two numbers is: ", number1 / number2)),
       "%/%" = print(paste("Integer Division of two numbers is: ", number1 %/% number2)),
       "%%" = print(paste("Division of two numbers is: ", number1 %% number2))
)
```

```{r}
number1<-readline(prompt = "enter number1:")
number2<-readline(prompt = "enter number2:")
number1=as.integer(number1)
number2=as.integer(number2)

operator <- readline(prompt="Please enter any ARITHMETIC OPERATOR You wish!: ")
 
switch(operator,
       "+" = print(paste("Addition of two numbers is: ", number1 + number2)),
       "-" = print(paste("Subtraction of two numbers is: ", number1 - number2)),
       "*" = print(paste("Multiplication of two numbers is: ", number1 * number2)),
       "^" = print(paste("Exponent of two numbers is: ", number1 ^ number2)),
       "/" = print(paste("Division of two numbers is: ", number1 / number2)),
       "%/%" = print(paste("Integer Division of two numbers is: ", number1 %/% number2)),
       "%%" = print(paste("the Division of num1 by num2 is: ", number1 %% number2))
)
```

```{r}
n1<-10
n2<-13
oper<-readline(prompt = "enter the operator:")
switch(oper,"+"=print(paste("addition is:",n1+n2)),"-"=print(paste("sub is:",n2-n1)),"/"=print(paste("division is",n1/n2)) )
```

```{r}
x<-letters[1:5]
for(i in x){
  print(i)
}
y<-LETTERS[1:9]
for (i in y){
  print(i)
}

```

```{r}
 
values <-  c(906, 264, 689, 739, 938)
 
barplot(values,col='red')
```



```{r}
cust[[5]]# to print column 5
cust[7,3]# 3 is the column, 7 is the 7th position in that clomn
```
```{r}
cust[c(1,2,3),c(2,3,4)]# to print the 1,2,3 rows and columns 2,3,4
cust$FirstName
cust[5,3]# accessing the element corresponding to row 5 and column 3
```
```{r}
head(cust, 3)# to print the first 3 rows
names(cust)# to print the names
tail(cust, 5) # to display the last 5 rows)

```

```{r}
summary(cust)
str(cust)
```

```{r}
barplot(cust$YearlyIncome,col="green",border = "red")
barplot(cust$YearlyIncome,col="green",border = "red",horiz = T,density =20)
cols = c("Black", "Blue", "brown","green", "Red", "gray", "White", "Yellow","orange","violet","Purple","Pink","Grey")
barplot(cust$YearlyIncome,col=cols,border="green")
```

```{r}
n3<-1:12
for (i in n3){
  print(2*i)
}

country<-c("ghana","togo","niger","benin")
for (a in country){
  print(paste("the country is:", a))
  }
```

```{r}
# FUNCTIONS
fun<-function(x1=4,x2=5,x3=6){
  m<-x1+x2+x3
  return(m)
}
fun(2,2,2)
fun()
```

```{r}
m_sum<-function(x){
  b=0
  for (i in seq(length(x))){
    b=b+x[i]
  }
    return(b)
}
m_sum(c(1,2,20,34,5,48,89))

pr<-function(y){
  h=1
  for (j in seq(length(y))){
  
    h=h*y[j]
  }
  return(h)
}
pr(c(1,2,3,2))
```




```{r}

su<-function(a,b,c)
  {
  a<-readline(prompt="enter a:")
b<-readline(prompt = "enter b:")
c<-readline(prompt = "enter c:")
a=as.integer(a)
b=as.integer(b)
c=as.integer(c)
  sum<-a+b+c
  av=sum/3
  print(paste("sum of",a,",",b,",",c,",","is=",sum))
    print(paste("av. of ",a,",",b,",",c,",","is=",av))
}
su(a,b,c)
```

```{r}
su<-function(a,b,c)
  {
  a<-readline(prompt="enter a:")
b<-readline(prompt = "enter b:")
c<-readline(prompt = "enter c:")
a=as.integer(a)
b=as.integer(b)
c=as.integer(c)
  sum<-a+b+c
  av=sum/3
  print(paste("sum of",a,",",b,",",c,",","is=",sum))
    print(paste("av. of ",a,",",b,",",c,",","is=",av))
}
```


```{r}
height <- c(132,151,162,139,166,147,122)
weight <- c(48,49,66,53,67,52,40)
gender <- c("male","male","female","female","male","female","male")
# Create the data frame.
input_data <- data.frame(height,weight,gender)
print(input_data)
# Test if the gender column is a factor.
print(is.factor(input_data$gender))
# Print the gender column so see the levels.
print(input_data$gender)
```
runif(1)
#> [1] 0.09006613

# Get a vector of 4 numbers
runif(4)
#> [1] 0.6972299 0.9505426 0.8297167 0.9779939

# Get a vector of 3 numbers from 0 to 100
runif(3, min=0, max=100)
#> [1] 83.702278  3.062253  5.388360

# Get 3 integers from 0 to 100
# Use max=101 because it will never actually equal 101
floor(runif(3, min=0, max=101))
#> [1] 11 67  1

# This will do the same thing
sample(1:100, 3, replace=TRUE)
#> [1]  8 63 64

# To generate integers WITHOUT replacement:
sample(1:100, 3, replace=FALSE)
#> [1] 76 25 52
To generate numbers from a normal distribution, use rnorm(). By default the mean is 0 and the standard deviation is 1.

rnorm(4)
#> [1] -2.3308287 -0.9073857 -0.7638332 -0.2193786

# Use a different mean and standard deviation
rnorm(4, mean=50, sd=10)
#> [1] 59.20927 40.12440 44.58840 41.97056

# To check that the distribution looks right, make a histogram of the numbers
x <- rnorm(400, mean=50, sd=10)
hist(x)
```{r}
getwd()
setwd("/home/emmanuelarize/BigData/Data/assignment3")
getwd()

a<-read.csv("/home/emmanuelarize/BigData/Data/assignment3/age.clean",header = T,sep = ',')
#a
b<-read.csv("/home/emmanuelarize/BigData/Data/assignment3/custdata.employed.clean",header = T,sep = ',')
#b
c<-read.csv("/home/emmanuelarize/BigData/Data/assignment3/num.vehicles.clean",header = T,sep=',')
#c
d<-read.csv("/home/emmanuelarize/BigData/Data/assignment3/housing.type.clean",header = T,sep = ',')
#d

#f<-cbind(a,b)# f and h are the same
#f

h<-data.frame(a,b)
g<-cbind(c,d)
g[1:2,]# to print the first two columns
```

```{r}
# generating uniform random numbers
runif(1)
#> [1] 0.09006613

# Get a vector of 4 numbers
runif(4)
#> [1] 0.6972299 0.9505426 0.8297167 0.9779939

# Get a vector of 3 numbers from 0 to 100
runif(3, min=0, max=100)
#> [1] 83.702278  3.062253  5.388360

# Get 3 integers from 0 to 100
# Use max=101 because it will never actually equal 101
floor(runif(3, min=0, max=101))
#> [1] 11 67  1

# This will do the same thing
sample(1:100, 3, replace=TRUE)
sample(1:13,3 ,replace = T)

#> [1]  8 63 64

# To generate integers WITHOUT replacement:
sample(1:100, 3, replace=FALSE)
sample(1:3,3)# by default is sample without replacement
#> [1] 76 25 52

# RANDOM NUMBERS FROM NORMAL DISTRIBUTION

#To generate numbers from a normal distribution, use rnorm().
#By default the mean is 0 and the standard deviation is 1.

rnorm(4)
#> [1] -2.3308287 -0.9073857 -0.7638332 -0.2193786

# Use a different mean and standard deviation
rnorm(4, mean=50, sd=10)
#> [1] 59.20927 40.12440 44.58840 41.97056

# To check that the distribution looks right, make a histogram of the numbers
x <- rnorm(400, mean=50, sd=10)
hist(x)
#par(mfrow=c(1,3))
```

```{python}
## installing a package in python
python -m pip install package_name

```

